---
title: "Homework 2"
author: "Felipe Dessoy Caraballo"
date: "02/05/2020"
output:
  html_document: default
  pdf_document: default
---

#Question 1A

```{r, echo=TRUE, warning=FALSE, message=FALSE}
rawData <- read.csv("https://raw.githubusercontent.com/FjjDessoyCaraballo/LU_QA_2020/master/Datasets/Corona_mergedV-Dem.csv")

library(ggplot2)
library(GGally)

ggplot(rawData, aes(x=cases_log))+
  geom_density(col='grey')

ggplot(rawData, aes(x=v2x_polyarchy))+
  geom_density(col='black')

ggplot(rawData, aes(x=v2x_libdem))+
  geom_density(col='blue')

ggplot(rawData, aes(x=v2x_partipdem))+
  geom_density(col='red')
```

#Question 1B

```{r, echo=TRUE, warning=FALSE, message=FALSE}

rawData$polyarchy_cat=ifelse(rawData$v2x_polyarchy> .5, "Democracy", "Autocracy")
rawData$polyarchy_cat<-factor(rawData$polyarchy_cat)
ggpairs(rawData, columns = c("popdata2018_log", "deaths_log", "v2x_partipdem", "polyarchy_cat"))

```

#Question 1C

```{r, echo=TRUE, warning=FALSE, message=FALSE}

#Polyarchy

rawData$polyarchy_cat=ifelse(rawData$v2x_polyarchy> .5, "Polyarchy", "Autocracy")
rawData$polyarchy_cat<-factor(rawData$polyarchy_cat)
boxplot(rawData$cases_log~rawData$polyarchy_cat, 
        col='yellow', xlab='Type of Democracy', 
        ylab='Number of Cases')

#Liberal Democracy

rawData$libdem_cat=ifelse(rawData$v2x_libdem > .5, "Liberal Democracy", "Autocracy")
rawData$libdem_cat<-factor(rawData$libdem_cat)
boxplot(rawData$cases_log~rawData$libdem_cat,
        col='blue',
        xlab='Type of Democracy', ylab='Number of Cases')

#Participatory Democracy

rawData$v2x_partipdem_cat=ifelse(rawData$v2x_partipdem > .5, "Participatory Democracy", "Autocracy")
rawData$v2x_partipdem_cat<-factor(rawData$v2x_partipdem_cat)
boxplot(rawData$cases_log~rawData$v2x_partipdem_cat,
        col='pink',
        xlab='Type of Democracy', ylab='Number of Cases')

```

#Question 1D1

```{r, echo=TRUE, warning=FALSE, message=FALSE}
OLS_m1=lm(cases_log~v2x_polyarchy, data=rawData)
summary(OLS_m1)
```

So far, our standar error is low and the regression line is a fit. That said, our variables here show us a positive correlation between polyarchy and the number of COVID-19 cases. However, this correlation might not be causal.

#Question 1D2

```{r,echo=TRUE, warning=FALSE, message=FALSE}
OLS_m2=lm(cases_log~v2x_libdem, data=rawData)
summary(OLS_m2)
```

Exchanging the polyarchy variable to the liberal democracy variable, we see an even lower standard error of 0.16, and our regression line has shown again a positive correlation to both independent and dependent variable. However, the causality link is still missing without a proper R².

#Question 1D3

```{r,echo=TRUE, warning=FALSE, message=FALSE}
OLS_m3=lm(cases_log~v2x_partipdem, data=rawData)
summary(OLS_m3)
```

The results of the 'participatory democracy' has shown similar results to the 'polyarchy' and 'liberal democracy' variables. Again, we have a low and acceptable standard error of 0.17 and our regression has proven to be positive. 

#Question 1D4

```{r,echo=TRUE, warning=FALSE, message=FALSE}
OLS_m4=lm(cases_log~v2x_polyarchy+v2x_libdem+v2x_partipdem, data=rawData)
summary(OLS_m4)

OLS_m5=lm(cases_log~v2x_polyarchy+popdata2018, data=rawData)
summary(OLS_m5)

rawData$norm_case=rawData$cases_log/rawData$popdata2018_log

OLS_m6=lm(norm_case~v2x_polyarchy, data=rawData)
summary(OLS_m6)
```

To this point, our data was showing a positive correlation between the number of COVID-19 cases and the different types of democracy. However, when we combine those three and compare to the cases of COVID-19, our correlation has proven the be negative in this case. The standard error has not changed significantly, except when considering our dependent variables. The results might have been altered considering number of population between them.

Question 1D5

```{r,echo=TRUE, warning=FALSE, message=FALSE}
OLS_m7=lm(norm_case~v2x_polyarchy+popdata2018, data=rawData)
summary(OLS_m7)

plot(rawData$v2x_polyarchy+rawData$popdata2018,rawData$norm_case, data=rawData, xlab="Norm Cases", ylab="Polyarchy and Population")
abline(lm(rawData$v2x_polyarchy+rawData$popdata2018~rawData$norm_case, data = rawData), col = "blue")
```

Finally, the data has been assimilated to an analysis by dividing the population into the 'polyarchy' variable, our new dependent variable is the 'normalized cases' instead of using corona cases. The results also have an acceptable standard error of 0.26. Although, the plot shows follows in a near parallel to the 'y' axis, which we can consider it to be a lack of correlation between the variables of population and 'polyarchy'.
